The \tsc{r}-pronoun and postposition \tit{waar-mee} in Dutch

### 2020-04-27 presentation nanolab/presentation.tex
klaar met daarmee voor nu
nog wel naar pavel sturen.. die zinnetjes

>>>>

\begin{frame}{The data}

\exg. Zij schrijft de brief \tbf{met} een potlood.\\
 she writes the letter with a pencil\\
 `She writes the letter with a pencil.'

\exg. Zij koopt \tbf{dat}.\\
 she buys that\\
 `She buys that.'

\exg. * Zij schrijft de brief \tbf{met} \tbf{dat}.\\
 she writes the letter with that\\
 `She writes the letter with that.'

\exg. Zij schrijft de brief \tbf{daar} \tbf{mee}.\\
 she writes the letter there with\\
 `She writes the letter with that.'

\begin{frame}{Questions and answers}

\centering{
\tbf{met dat} | \tbf{daar mee} \pause
}

\begin{enumerate}
  \item Q: why *\tit{met dat} (but \tit{daar mee})? \pause\\
  A: \tit{daar mee} is the default, \tit{met dat} only appears under special circumstances (morphological derivation, cf. \tit{mice} and not \tit{mouse-s})\pause
  \item Q: why does the adposition change form? \pause\\
  A: \tit{mee} and \tit{met} are stored as different forms in the lexicon \pause
  \item Q: why does the adposition change position? \pause\\
  A: \tit{mee} and \tit{met} are stored as preposition or postposition in the lexicon \pause
  \item Q: why is there an \tit{r}, even though there is no location involved? \pause\\
  A: \tit{r} is ambiguous between \tsc{thing} and \tsc{location}
\end{enumerate}

\begin{frame}{Nanosyntax}

\begin{itemize}
  \item functional sequence is constant\pause
  \item spellout algorithm is constant\pause
  \item a single place for cross-linguistic differences: the postsyntactic lexicon\pause
\end{itemize}

\begin{enumerate}
  \item spellout happens feature by feature and targets phrases
  \item phonological forms are stored in the postsyntactic lexicon
  \item postpositions are candidates for spellout before prepositions are
  \item ambiguity comes from containment relations
\end{enumerate}

\begin{frame}{differing phonological forms}

\begin{itemize}
  \item \tit{mee} and \tit{met} are stored as different forms in the lexicon
  \item phonological forms are stored in the lexicon \pause
\end{itemize}

\begin{frame}{pre vs. post}

\begin{itemize}
  \item \tit{mee} and \tit{met} are stored as preposition or postposition in the lexicon \pause
  \item postpositions are candidates for spellout before prepositions are \pause
\end{itemize}

\begin{frame}{pre vs. post}

\begin{frame}{\tit{r} without location}

  \begin{itemize}
    \item \tit{r} is ambiguous between \tsc{thing} and \tsc{location}
    \item ambiguity comes from containment relations
  \end{itemize}

\begin{frame}{Derivational system}

\begin{itemize}
  \item \tit{daar mee} is the default, \tit{met dat} only appears under special circumstances (morphological derivation, cf. \tit{mice} and not \tit{mouse-s}) \pause
  \item spellout happens feature by feature and targets constituents
\end{itemize}

\ex.
\ag. Ze heeft de brief \tbf{daar} \tbf{mee} geschreven.\\
 she has the letter there with written\\
 `She wrote the letter with that.'
\bg. Zij heeft de brief \tbf{met} \tbf{dat} potlood geschreven.\\
 she has the letter with that pencil written\\
 `She wrote the letter with that potlood.'

\begin{frame}{\tit{daar mee} in a proper constituent}

\exg. Ze heeft de brief \tbf{daar} \tbf{mee} geschreven.\\
 she has the letter there with written\\
 `She wrote (the letter) with that.'

\exg. \tbf{Daar} \tbf{mee} heeft ze de brief geschreven.\\
 there with has she the letter written\\
 `With that she has written the letter.' \hfill topicalization

\exg. Ze heeft de brief \tbf{hier} \tbf{mee} en \tbf{daar} \tbf{mee} geschreven.\\
 she has the letter here with and there with written\\
 `She has written the letter with this and with that' \hfill coordination

\ex.
\a. With what did she write the letter?
\bg. \tbf{daar} \tbf{mee}\\
 there with\\
 `with that' \hfill fragment answer

  \ex. \begin{forest} boom
  [\tsc{insP}
      [\tsc{f4}]
  	    [\tsc{datP}
  	        [\tsc{f3}]
  	        [\tsc{accP}
  	           [\tsc{f2}]
  	           [\tsc{nomP}
  	               [\tsc{f1}]
  	               [ΣP
  	                   [Σ]
  										 [IP
  										 		[I]
  			                   [\tsc{thingP}
  			                       [\tsc{thing}, roof]
  												 ]
  	                   ]
  	               ]
  	           ]
  	        ]
  	    ]
  	    {\draw (.east) node[right]{⇒ \tit{'t}}; }
    ]
  \end{forest}

\begin{frame}{\tit{met dat} in a non-constituent}

\exg. Ze heeft de brief \tbf{met} \tbf{dat} potlood geschreven.\\
she has the letter with the pencil written\\
`She has written the letter with that pencil.'

\ex.
\ag. *\tbf{Met} \tbf{dat} heeft ze de brief \tbf{potlood} geschreven.\\
 with that has she the letter pencil written\\
 `With that pencil she has written the letter.'
\bg. \tbf{Met} \tbf{dat} \tbf{potlood} heeft ze de brief geschreven.\\
 with that pencil has she the letter written\\
 `With that pencil she has written the letter.' \hfill topicalization

% \ex.
% \ag. *Ze heeft de brief met die en met dat potlood geschreven.\\
%  she has the letter with this and with that pencil written\\
%  `She has written the letter with this pencil and with that marker.' \hfill coordination
% \bg. Ze heeft de brief met die pen en en dat potlood geschreven.\\
%  she has the letter with this marker and with that pencil written\\
%  `She has written the letter with this pencil and with that marker.' \hfill coordination
%
% \pause

\ex.
\a. With what did she write the letter?
\bg. *\tbf{met} \tbf{dat}\\
 with that\\
 `with that'
\bg. \tbf{met} \tbf{dat} \tbf{potlood}\\
 with that pancil\\
 `with that pencil'\hfill fragment answer

  \ex. \begin{forest} boom
[\tsc{insP}
    [\tsc{f4}]
    [\tsc{datP}
        [\tsc{datP}
            [\tsc{f3}]
            [\tsc{accP}
               [\tsc{f2}]
               [\tsc{nomP}
                   [\tsc{f1}]
                   [ΣP
                       [Σ]
  										 [IP
  										 		[I]
                           [\tsc{thingP}
                               [\tsc{thing}, roof]
                           ]
  											]
                   ]
               ]
            ]
        ]
        {\draw (.east) node[right]{⇒ \tit{t}}; }
        [NP
            [\tsc{pencil}, roof]
        ]
        {\draw (.east) node[right]{⇒ \tit{potlood}}; }
    ]
]
    \end{forest}

\begin{frame}{Summary}

    \begin{enumerate}
      \item Q: why *\tit{met dat} (but \tit{daar mee})? \pause\\
      A: \tit{daar mee} is the default, \tit{met dat} only appears under special circumstances (morphological derivation, cf. \tit{mice} and not \tit{mouse-s})\pause
      \item Q: why does the adposition change form? \pause\\
      A: \tit{mee} and \tit{met} are stored as different forms in the lexicon \pause
      \item Q: why does the adposition change position? \pause\\
      A: \tit{mee} and \tit{met} are stored as preposition or postposition in the lexicon \pause
      \item Q: why is there an \tit{r}, even though there is no location involved? \pause\\
      A: \tit{r} is ambiguous between \tsc{thing} and \tsc{location}
    \end{enumerate}
